---
title: "Coursera Machine Learning"
output: html_document
---

# Practical Machine Learning: Course Project

## Objective
The goal of the project is to build a machine learning algorithm in order to predict a variable Classe from a database containing a variety of indicators concerning body movements, gathered by devices such as Fitbit.

## Cleaning and ordering data
Before we start building models we have to clean our data from all the columns with NAs and eliminate first seven variables with no predictive value.

```{r}
library(randomForest)
library(caret)


training <- read.csv("C:\\Data\\R\\Machine learning\\pml-training.csv", na.strings=c("NA", ""))
training <-training[,colSums(is.na(training)) == 0]
testing <- read.csv("C:\\Data\\R\\Machine learning\\pml-testing.csv", na.strings=c("NA", ""))
testing <-testing[,colSums(is.na(training)) == 0]
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
```

Since our testing set is very small, and the training one very big, we should divide our training set into two to verify its validity: Train1 for training and Train2 for testing. 

```{r}
Train <- createDataPartition(training$classe, p = 0.6, list=FALSE)
Train1 <- training[Train,]
Train2 <- training[-Train,]

```

## Models
Two models will be explored: rpart and random forest.
Let us have a look on rpart and test on our Train2 data

```{r}
rpartTrain <- train(classe ~ ., data = Train1, method="rpart")
predictions <- predict(rpartTrain, newdata=Train2)
confusionMatrix(predictions, Train2$classe)
```
The accuracy is very low, only about 0.5.

Let us try random forest then.
```{r}
rfTrain <- randomForest(classe ~. , data=Train1)
predictions <- predict(rfTrain, newdata=Train2)
confusionMatrix(predictions, Train2$classe)
```
In this case accuracy is very high, almost 1. Random forest seems to be a preferable model.

#Conclusions
Having discovered that random forest is a better choice (as indicated in lectures), we can apply our model to the definite testing data
```{r}
predict(rfTrain, newdata=testing)
```
It gives correct results.